Quantifying uncertainty in Pareto estimates of global lake area

### 2021-10-20 manuscript/manuscript.tex
minor text edits

\begin{abstract}
	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration and greenhouse gas emissions. Because we do not have a complete census of all lakes, upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area. Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to by omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e. "censoring" of small lakes relative to large lakes).

	The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions. Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point). This yields an over-exact lake area estimate that is overwhelmingly reported with no uncertainty bounds. I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates. I show that such although such models produce a more realistic estimate of lake area uncertainty they underestimate true total lake area. The degree of this underestimation is likely related to the proportion of the dataset subject to censoring. Ultimately, this may explain the fact that total lake area estimates have increased through time as the resolution of lake databases has improved.
\end{abstract}

>>>>

\begin{abstract}
	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration and greenhouse gas emissions. Because we do not have a complete census of all lakes, upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area. Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to be omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e. "censoring" of small lakes relative to large lakes). The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions. Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point). This yields an over-exact lake area estimate that is overwhelmingly reported with no uncertainty bounds. I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates. I show that such although such models produce a more realistic estimate of lake area uncertainty they underestimate true total lake area. The degree of this underestimation is likely related to the proportion of the dataset subject to censoring. Ultimately, this may explain the fact that total lake area estimates have increased through time as the resolution of lake databases has improved.
\end{abstract}
